---
title: "Data 605 Assignment 6"
author: "Tyler Baker"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. A bag contains 5 green and 7 red jellybeans. How many ways can 5 jellybeans be
withdrawn from the bag so that the number of green ones withdrawn will be less than
2?

Ans: There are 2 possibilities:
5 red, 0 green
4 red, 1 green
```{r}
case_a <- choose(7,5) * choose(5,0)
case_b <- choose(7,4) * choose(5,1)
print(case_b + case_a)
```

2. A certain congressional committee consists of 14 senators and 13 representatives. How
many ways can a subcommittee of 5 be formed if at least 4 of the members must be
representatives?

Ans: There are 2 possibilities:
4 representatives, 1 senator
5 representaties, 0 senators
```{r}
case_a <- choose(13, 4) * choose(14,1)
case_b <- choose(13,5) * choose(14,0)
print(case_a + case_b)
```

3. . If a coin is tossed 5 times, and then a standard six-sided die is rolled 2 times, and finally
a group of three cards are drawn from a standard deck of 52 cards without
replacement, how many different outcomes are possible?

Ans:
```{r}
coin_tosses <- 2^5
dice_rolls <- 6^2
cards_drawn <- 52 * 51 * 50
print(coin_tosses + dice_rolls + cards_drawn)
```

4. 3 cards are drawn from a standard deck without replacement. What is the probability
that at least one of the cards drawn is a 3? Express your answer as a fraction or a
decimal number rounded to four decimal places.


Ans: A/B
B = Total number of ways to draw 3 cards without replacement.
A = # of ways the 3 cards have at least one three.
For A there are a few possibilities:
case1 = (3, blank, blank) , (blank, 3, blank) , (blank, blank, 3)
case2 = (3, 3, blank), (3, blank, 3), (blank, 3, 3)
case3 = (3, 3, 3)
```{r}
b <- 52 * 51 * 50
a <- 7
print(a/b)
```

5. Lorenzo is picking out some movies to rent, and he is primarily interested in
documentaries and mysteries. He has narrowed down his selections to 17
documentaries and 14 mysteries.

a.  How many different combinations of 5 movies can he rent?

Ans: There's 31 total movies. Therefore,
```{r}
print(31 * 30 * 29 * 28 * 27)
```

b. How many different combinations of 5 movies can he rent if he wants at least
one mystery?

Ans: There are a few cases to consider.
case1: there is one mystery
case2: there is two mystery movies
case3: there is three mystery movies
case4: there is four mystery movies
case5: there is five mystery movies
```{r}
case_a <- choose(14, 1) * choose(17, 4)
case_b <- choose(14, 2) * choose(17, 3)
case_c <- choose(14,3) * choose(17,2)
case_d <- choose(14,4) * choose(17,1)
case_e <- choose(14,5) * choose(17,0)
print(case_a + case_b + case_c + case_d + case_e)
```

6. In choosing what music to play at a charity fund raising event, Cory needs to have an
equal number of symphonies from Brahms, Haydn, and Mendelssohn. If he is setting up
a schedule of the 9 symphonies to be played, and he has 4 Brahms, 104 Haydn, and 17
Mendelssohn symphonies from which to choose, how many different schedules are
possible? Express your answer in scientific notation rounding to the hundredths place.

Ans: Since there are 9 songs and Brahms, Haydn, Mendelssohn needs to be equal amounts, they each need 3 songs played.
```{r}
brahms <- choose(4,3)
haydn <- choose(104,3)
mendelssohn <- choose(17,4)
print(brahms + haydn + mendelssohn)
```

7. An English teacher needs to pick 13 books to put on his reading list for the next school
year, and he needs to plan the order in which they should be read. He has narrowed
down his choices to 6 novels, 6 plays, 7 poetry books, and 5 nonfiction books.

a. If he wants to include no more than 4 nonfiction books, how many different
reading schedules are possible? Express your answer in scientific notation rounding to
the hundredths place. 

Ans: There are a few cases to consider.
case1: there are 0 nonfiction books.
case2: there is 1 nonfiction book.
case3: there are 2 nonfiction books.
case4: there are 3 nonfiction books.
case5: there are 4 nonfiction books.

Since, there are no parameters on the other genres we can say that any of them are possibilities after the nonfiction picks.
```{r}
case_a <- choose(5,0) + (16*15*14*13*12*11*10*9*8*7*6*5*4)
case_b <- choose(5,1) + (16*15*14*13*12*11*10*9*8*7*6*5)
case_c <- choose(5,2) + (16*15*14*13*12*11*10*9*8*7*6)
case_d <- choose(5,3) + (16*15*14*13*12*11*10*9*8*7)
case_e <- choose(5,4) + (16*15*14*13*12*11*10*9*8)
print(case_a + case_b + case_c + case_d + case_e)
```

b. If he wants to include all 6 plays, how many different reading schedules are
possible? Express your answer in scientific notation rounding to the hundredths place.

Ans: There will be 13 spaces where 6 will have to be filled.

After finding out how many ways to put 6 objects into 13 spaces, we can then consider the cases from above.
Which is choose(13,6)

Note that there will only be 10 books left.

```{r}
case_a <- choose(13,6) + choose(5,0) + (13*12*11*10*9*8*7)
case_b <- choose(13,6) + choose(5,1) + (13*12*11*10*9*8)
case_c <- choose(13,6) + choose(5,2) + (13*12*11*10*9)
case_d <- choose(13,6) + choose(5,3) + (13*12*11*10)
case_e <- choose(13,6) + choose(5,4) + 
  (13*12*11)
print(case_a  +case_b + case_c + case_d + case_e)
```

8. Zane is planting trees along his driveway, and he has 5 sycamores and 5 cypress trees to
plant in one row. What is the probability that he randomly plants the trees so that all 5
sycamores are next to each other and all 5 cypress trees are next to each other?
Express your answer as a fraction or a decimal number rounded to four decimal places.

Ans: There are 10 spots total and there are 5 of each tree. The answer will be the # of ways the condition is satisfied divided by the total number of ways the trees could be planted.

There are only two ways that the condition is met. Either all of the sycamores are planted first or all of the cypress are planted first.

The total possible ways to plant the trees is 2^10. This is because each spot could be either a sycamore or a cypress.
```{r}
print(2/(2^10))
```

9. If you draw a queen or lower from a standard deck of cards, I will pay you $4. If not, you
pay me $16. (Aces are considered the highest card in the deck.)

P(losing) = 8/52
P(winning) = 44/52

a.  Find the expected value of the proposition. Round your answer to two decimal
places. Losses must be expressed as negative values

```{r}
p_winning <- 44/52
p_losing <- 8/52

e_v <- (p_winning* 4) - (p_losing*16) 
```


b. If you played this game 833 times how much would you expect to win or lose?
Round your answer to two decimal places. Losses must be expressed as negative
values.

```{r}
e_v * 833
```

